```{r setoptions, echo = FALSE}
library(knitr)
library(plyr)
library(ggplot2)
opts_chunk$set(echo = TRUE)
```
---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the "quantified self" movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

## Loading and preprocessing the data
1. The process verifies if the zip file with the collected data has already been downloaded and, if not, it gets it from the specified URL.
2. The file is unzipped to get the CSV file.
3. The process loads the data set in a variable called *activity*. The NA values are asigned using the "NA" string and the column "date" is formated as date data.

```{r}
if(!file.exists("activity.zip")) {
  download.file(
    "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",
    "activity.zip")
}

if(!file.exists("activity.csv")) {
  unzip("activity.zip")
}

activity <- read.csv("activity.csv", header = TRUE, na.strings = "NA")

activity[,2] <- as.Date(activity[,2], "%Y-%m-%d")
```


## What is mean total number of steps taken per day?

The total number of steps taken per day is represented as follows:

```{r}

totalSteps <- ddply(
              activity, 
              ~date, 
              summarise, 
              totalsteps = sum(steps, na.rm = TRUE))

qplot(
  date,
  weight = totalsteps,
  data = totalSteps,
  xlab = "Date",
  ylab = "Steps",
  main = "Total number of steps taken per day",
  fill=I("grey"),
  col=I("black"),
  binwidth = 0.5,
  geom="histogram")
```

And here we have the mean and median of the total number of steps taken per day:

```{r}
stepMeanByDate <- ddply(
              na.omit(activity), 
              ~date, 
              summarise, 
              mean = mean(steps, na.rm = TRUE),
              median = median(steps, na.rm = TRUE))
names(stepMeanByDate) <- c("Date", "Step Mean", "Median Steps")
print(stepMeanByDate)
```

## What is the average daily activity pattern?

```{r}
stepMeanByInt <- ddply(
                  activity, 
                  .(interval), 
                  summarise, 
                  mean = mean(steps, na.rm = TRUE))

qplot(
  		interval, 
			y = mean, 
			data = stepMeanByInt, 
			xlab = "Interval", 
			ylab = "Steps",
			main = "Average daily activity pattern",
			geom = "line")

```

And we can calculate the max interval row:

```{r}
stepMeanByInt[which.max(stepMeanByInt[,2]),]
```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
